---
title: "Vaccination"
output:
  rmarkdown::html_document
    
---

```{r global_options}
knitr::opts_chunk$set(fig.path='Figs/')
```

### Import data & library

### setup wd
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/2021/Vacinação/Data/EXCEL")
```

### read.csv
```{r}
setwd("~/2021/Vacinação/Data/EXCEL")
covid2<-read.table("covid_15_03.csv",sep = ",")
```

### packages
```{r, results="hide", warning=FALSE, message=FALSE}
packages <- c("dplyr", "tibble", "tidyr", "purrr", "FactoMineR", "ggplot2", "lm.beta", "olsrr", "neuralnet", "nnet", "caret", "NeuralNetTools", "reticulate","factoextra","lavaan", "FactoMineR", "lavaan", "extrafont", "rcompanion", "psy")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)
``` 

## Exploratory analysis

### Graph Vaccination Intention
```{r, warning=FALSE, message=FALSE}
covid2[,60:117] <- as.numeric((unlist(covid2[,60:117])))

# NA removal
covid2 <- covid2[!is.na(covid2$V87), ]
covid2 <- covid2[!is.na(covid2$V60), ]
covid2 <- covid2[!covid2$V23==17, ]
covid2 <- covid2[!is.na(covid2$V88), ]

# children vaccination intention
covid2[,56:59] <- as.numeric((unlist(covid2[,56:59])))

covid2%>%
   mutate(intention_children = rowSums(cbind(V57, V58, V59), na.rm=T)) ->covid2

# vaccination col to long format
longdata2 = covid2 %>%
   gather("V56", "intention_children", key = Vaccination, value = Intention)

# rename value own & children
longdata2$Vaccination[longdata2$Vaccination=="V56"]<-"self"
longdata2$Vaccination[longdata2$Vaccination=="intention_children"]<-"children" 

# group by Intention & Vaccination
longdata2 %>%
  dplyr::group_by(Intention, Vaccination) %>%
summarise(Frequency = dplyr::n()) -> graph_mutate2
graph_mutate2$Frequency <- graph_mutate2$Frequency/649*100

# graph vaccination intention
graph_mutate2 %>% 
   mutate(Vaccination = as.factor(Vaccination)) %>% 
   ggplot(aes(x = Intention, y=Frequency, fill = Vaccination)) + 
  geom_bar(stat="identity", position=position_dodge()) + labs(x = "Intention to vaccinate", y= " Relative frequency (%)") +
   scale_x_discrete(expand = c(0, 0), name ="Intention of having the COVID-19 vaccine", 
                    limits=c("1","2","3","4","5","6","7")) +  geom_text(aes(label=sprintf("%1.1f%%", Frequency)), position = position_dodge(0.9),
                                                                        vjust = -0.1, size = 3) + scale_y_continuous(expand = c(0, 0), breaks = c(0,10,20,30,40,50,60,70))+
  coord_cartesian(ylim=c(0, 70)) +  theme_bw() +
  theme_minimal() +  theme(panel.grid.minor.y = element_blank(),
panel.grid.minor.x = element_blank(),panel.grid.major.x = element_blank() ) + theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                                               axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid')) +
theme(text=element_text(family="Times New Roman", size=12)) +  theme(axis.text = element_text(size = 12),
    axis.text.y = element_text(color = "black"), axis.text.x = element_text(color = "black")) + theme(plot.margin = unit(c(1,3,1,1), "cm")) + scale_fill_grey() + theme(panel.grid.major = element_blank())

```

### Graph Vaccination Concerns
```{r, warning=FALSE, message=FALSE}
covid2[,118:123]<- as.numeric(unlist(covid2[,118:123]))

# matrix concerns_vaccine
concerns_vaccine <- matrix(c(sum(covid2$V119, na.rm = T), sum(covid2$V118, na.rm = T), sum(covid2$V120, na.rm = T), sum(covid2$V121, na.rm = T), sum(covid2$V122, na.rm = T)), ncol=5, byrow=TRUE)

# data_concerns to data.frame 
data_concerns <-as.data.frame(concerns_vaccine)

# concerns to long format
longdata4 = data_concerns %>%
  gather("V1", "V2", "V3","V4","V5", key = Concerns, value = Frequency)

longdata4$Frequency <- round(longdata4$Frequency/649*100, digits = 1)

# revalue concerns
longdata4$Concerns[longdata4$Concerns=="V1"]<-"record time"
longdata4$Concerns[longdata4$Concerns=="V2"]<-"side effects"
longdata4$Concerns[longdata4$Concerns=="V3"]<-"risk group"
longdata4$Concerns[longdata4$Concerns=="V4"]<-"ineffective"
longdata4$Concerns[longdata4$Concerns=="V5"]<-"microchip"

#concerns as factor 
longdata4$Concerns<-as.factor(longdata4$Concerns)
longdata4$Concerns <- factor(c("record time", "side effects", "risk group", "ineffective", "microchip"),
                levels = c("record time", "side effects", "risk group", "ineffective", "microchip"))

# concerns graph
ggplot(data=longdata4, aes(x=Concerns, y=Frequency)) +
  geom_bar(stat="identity")+ coord_cartesian(ylim = c(0, 70))+ 
  theme_minimal()  + theme(panel.grid.minor = element_blank()) +
scale_y_continuous(expand = c(0, 0), breaks = c(0,10,20,30,40,50,60,70))+
  coord_cartesian(ylim=c(0, 70)) + theme(panel.grid.major.y = element_blank(),
panel.grid.minor.y = element_blank() ) + theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                                               axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid')) +
  geom_text(aes(label=sprintf("%1.1f%%", Frequency)), position = position_dodge(0.9),
                                                                        vjust = -0.1, size = 3) + labs(x = "Concerns", y= " Relative frequency (%)") +
  scale_x_discrete(expand = c(0, 0)) + theme(plot.margin = unit(c(5, 5, 5, 5), "mm")) + 
  theme(text=element_text(family="Times New Roman", size=12)) + theme(panel.grid.major = element_blank())


```

### Coding variables
```{r, warning=FALSE, message=FALSE}
covid2$V2<-as.numeric(unlist(covid2$V2))

covid2$V21[covid2$V21==3]<-"male"
covid2$V21[covid2$V21==4]<-"female"
covid2$V21[covid2$V21==5]<- NA
covid2$V21<-as.factor(covid2$V21)

covid2$V23<-as.numeric(unlist(covid2$V23))

covid2$V25[covid2$V25==1]<-"aaprimary/basic"
covid2$V25[covid2$V25==2]<-"aaprimary/basic"
covid2$V25[covid2$V25==3]<-"aaprimary/basic"
covid2$V25[covid2$V25==4]<-"high school"
covid2$V25[covid2$V25==5]<-"academic"
covid2$V25[covid2$V25==6]<-"academic"
covid2$V25[covid2$V25==7]<-"academic"
covid2$V25[covid2$V25==8]<- NA
covid2$V25<-as.factor(covid2$V25)

covid2$V27[covid2$V27==1]<-"rural"
covid2$V27[covid2$V27==2]<-"urban"
covid2$V27<-as.factor(covid2$V27)

covid2$V28[covid2$V28==1]<-"yes"
covid2$V28[covid2$V28==2]<-"no"
covid2$V28<-as.factor(covid2$V28)

covid2$V29<-as.numeric(unlist(covid2$V29))

covid2$V30[covid2$V30==1]<-"working"
covid2$V30[covid2$V30==2]<-"unemployed"
covid2$V30[covid2$V30==3]<-"retired"
covid2$V30[covid2$V30==5]<-"student"
covid2$V30[covid2$V30==6]<-"working"
covid2$V30[covid2$V30==7]<- NA
covid2$V30<-as.factor(covid2$V30)

covid2$V32[covid2$V32==1]<-"yes"
covid2$V32[covid2$V32==2]<-"no"
covid2$V32<-as.factor(covid2$V32)

covid2$V48<-as.numeric(unlist(covid2$V48))

covid2$V49[covid2$V49==1]<-"yes"
covid2$V49[covid2$V49==2]<-"no"
covid2$V49<-as.factor(covid2$V49)

covid2$V51<-as.numeric(unlist(covid2$V51))

covid2$V54<-as.numeric(unlist(covid2$V54))

covid2$V55[covid2$V55==1]<-"no"
covid2$V55[covid2$V55==2]<-"no"
covid2$V55[covid2$V55==3]<-"yes"
covid2$V55[covid2$V55==4]<-"yes"
covid2$V55<-as.factor(covid2$V55)
```

### Dimensionality reduction - Scales 

### PCA COVID-19 scale 
```{r, warning=FALSE, message=FALSE}

pca <- PCA(covid2[,60:67], graph = F)

get_eig(pca)

fviz_screeplot(pca, addlabels = TRUE, ylim = c(0, 50))
var <- get_pca_var(pca)
fviz_pca_var(pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE
)
summary(pca,nbelements = 10, ncp = 3)
```

### CFA COVID-19 scale
```{r, warning=FALSE, message=FALSE}

model<-'
covid_threat=~V60 + V62 + V63 + V64
trust=~V66 + V67
covid_impact=~V65 + V61'

fit<- cfa(model, data=covid2)
summary(fit, fit.measures=T,standardized=T)
```

### coding COVID-19 scale dimensions
```{r, warning=FALSE, message=FALSE}
covid2 %>%
  rowwise %>%
  mutate(covid_threat = mean(c(V60, 8-V62, V63, 8-V64))) %>%
  mutate(trust = mean(c(V66, V67))) -> covid2 
  
covid2 %>%
  rowwise %>%
  mutate(covid_impact = mean(c(V65, 8-V61))) ->covid2
```

### alpha COVID-19 global
```{r, warning=FALSE, message=FALSE}
alpha_covid_threat <- covid2 %>% select(V60, V62, V63, V64, V66, V67, V65, V61)
alpha_covid_threat$V62<-8-alpha_covid_threat$V62
alpha_covid_threat$V64<-8-alpha_covid_threat$V64
alpha_covid_threat$V66<-8-alpha_covid_threat$V66
alpha_covid_threat$V67<-8-alpha_covid_threat$V67
alpha_covid_threat$V61<-8-alpha_covid_threat$V61

cronbach(alpha_covid_threat)
```

### alpha COVID-19 covid_threat
```{r, warning=FALSE, message=FALSE}
alpha_covid_threat <- covid2 %>% select(V60, V62, V63, V64)
alpha_covid_threat$V62<-8-alpha_covid_threat$V62
alpha_covid_threat$V64<-8-alpha_covid_threat$V64
cronbach(alpha_covid_threat)
```

### alpha COVID-19 covid_trust
```{r, warning=FALSE, message=FALSE}
alpha_covid_trust <- covid2 %>% select(V66, V67)

cronbach(alpha_covid_trust)
```
### alpha COVID-19 global
```{r, warning=FALSE, message=FALSE}
alpha_covid_impact <- covid2 %>% select(V65, V61)
alpha_covid_impact$V61<-8-alpha_covid_impact$V61


cronbach(alpha_covid_impact)
```

### PCA vaccine scale 
```{r, warning=FALSE, message=FALSE}
pca <- PCA(covid2[,68:87], graph=F)

get_eig(pca)

fviz_screeplot(pca, addlabels = TRUE, ylim = c(0, 50))
var <- get_pca_var(pca)

fviz_pca_var(pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE 
)

summary(pca,nbelements = 20, ncp = 4)
```

### CFA vaccine scale 
```{r, warning=FALSE, message=FALSE}

model<-'
beliefs_attitudes=~V68+V69+V70+V71+V72+V73+V74+V75+V76+V82+V83+V84+V85+V86
others_intention=~V77+V78
knowledge=~V80+V81
return_to_normal=~V79+V87'
fit<- cfa(model, data=covid2)
summary(fit, fit.measures=T,standardized=T)
```

### coding vaccine scale dimensions
```{r, warning=FALSE, message=FALSE}
covid2 %>%
  rowwise %>%
  mutate(beliefs_attitudes = mean(c(V68, V69, V70, V71, V72, 8-V73, 8-V74, 8-V75, 8-V76, V82, V83, V84, V85, V86))) %>%
  mutate(others_intention = mean(c(V77, V78))) %>%
  mutate(knowledge = mean(c(V80, V81))) %>%                                     
  mutate(return_to_normal = mean(c(V79, V87))) -> covid2 
```

### alpha COVID-19 vaccine global
```{r, warning=FALSE, message=FALSE}
alpha_covid_beliefs <- covid2 %>% select(V68, V69, V70, V71, V72, V73, V74, V75, V76, V82, V83, V84, V85, V86, V77, V78, V80, V81, V79, V87)
alpha_covid_beliefs$V73<-8-alpha_covid_beliefs$V73
alpha_covid_beliefs$V74<-8-alpha_covid_beliefs$V74
alpha_covid_beliefs$V75<-8-alpha_covid_beliefs$V75
alpha_covid_beliefs$V76<-8-alpha_covid_beliefs$V76
cronbach(alpha_covid_beliefs)
```

### alpha COVID-19 vaccine beliefs
```{r, warning=FALSE, message=FALSE}
alpha_covid_beliefs <- covid2 %>% select(V68, V69, V70, V71, V72, V73, V74, V75, V76, V82, V83, V84, V85, V86)
alpha_covid_beliefs$V73<-8-alpha_covid_beliefs$V73
alpha_covid_beliefs$V74<-8-alpha_covid_beliefs$V74
alpha_covid_beliefs$V75<-8-alpha_covid_beliefs$V75
alpha_covid_beliefs$V76<-8-alpha_covid_beliefs$V76
cronbach(alpha_covid_beliefs)
```
### alpha COVID-19 others_intention
```{r, warning=FALSE, message=FALSE}
alpha_covid_other <- covid2 %>% select(V77, V78)

cronbach(alpha_covid_other)
```
### alpha COVID-19 others_intention
```{r, warning=FALSE, message=FALSE}
alpha_covid_knowledge <- covid2 %>% select(V80, V81)

cronbach(alpha_covid_knowledge)
```
### alpha COVID-19 others_intention
```{r, warning=FALSE, message=FALSE}
alpha_covid_return_normal <- covid2 %>% select(V79, V87)

cronbach(alpha_covid_return_normal)
```

### coding risk aversion scale
```{r, warning=FALSE, message=FALSE}
covid2 %>%
  rowwise %>%
  mutate(risk_aversion = mean(c(V88, V89, V90, V91, V92, V93, V94, V95, V96, V97, V98, V99, V100, V101, V102,V103, V104,V105, V106, V107,V108,V109,V110, V111,V112, V113,V114,V115, V116,V117)))->covid2
```

### rename col
```{r, warning=FALSE, message=FALSE}
names(covid2)[2]  <- "days_lockdown"
names(covid2)[21] <- "gender"
names(covid2)[23] <- "age"
names(covid2)[24] <- "nationality"
names(covid2)[25] <- "literacy"
names(covid2)[26] <- "age"
names(covid2)[27] <- "residency"
names(covid2)[28] <- "relationship"
names(covid2)[29] <- "socioeconomic"
names(covid2)[30] <- "professional_status"
names(covid2)[32] <- "have_children"
names(covid2)[48] <- "how_much_life_changed"
names(covid2)[49] <- "religious"
names(covid2)[50] <- "religion"
names(covid2)[51] <- "economic_changes"
names(covid2)[54] <- "covid-19_risk"
names(covid2)[55] <- "had_covid-19"
names(covid2)[56] <- "vaccination_intention"

covid2$socioeconomic[covid2$socioeconomic==1]<-"low"
covid2$socioeconomic[covid2$socioeconomic==2]<-"low"
covid2$socioeconomic[covid2$socioeconomic==3]<-"1medium"
covid2$socioeconomic[covid2$socioeconomic==4]<-"high"
covid2$socioeconomic[covid2$socioeconomic==5]<-"high"
covid2$socioeconomic<-as.factor(covid2$socioeconomic)

covid2 <- covid2[!is.na(covid2$risk_aversion), ]
```

## LM models

### t-test children & own
```{r, warning=FALSE, message=FALSE}
# t-test own & children
t.test(covid2$vaccination_intention, covid2$intention_children, paired = TRUE, alternative = "two.sided")
```

### Parsimonious models (MLR)
### Own vaccination intention
```{r, warning=FALSE, message=FALSE}
modelintencao2 <- lm(covid2$vaccination_intention ~ covid2$`covid-19_risk` + covid2$knowledge +
covid2$beliefs_attitudes)
summary(modelintencao2)
lm.beta(modelintencao2)

```

### Children vaccination
```{r, warning=FALSE, message=FALSE}
modelfilhos2 <- lm(covid2$intention_children ~ covid2$covid_threat + covid2$knowledge + covid2$beliefs_attitudes)
summary(modelfilhos2)
lm.beta(modelfilhos2)

```
